Live and Let Die? (Delayed Issues not Harder to Resolve)

### 2015-08-20 data.tex
Updates from ShareLaTeX

Since 2000, the SEI has been teaching and coaching TSP teams. One of the authors (Nichols) has been been mentoring software development teams and coaches around the world as they deploy TSP within their organizations since 2006.  The  most recent completions were in 2014.
The projects were mostly small to medium, with a median duration of 46 days and a maximum duration of 90 days in major increments. 
Several projects extended for multiple incremental development cycles. 
Median team size was 7 people, with a maximum of 40. See \fig{dist} for the total
effort seen in those projects.

@@

Note that, in that Figure~\ref{fig:waterfall},
several  phases in which product is created have sub-phases of {\em review} and {\em inspect} to remove defects.
In TSP  reviews,  individuals perform personal reviews of their work products prior to the peer review
(which TSP calls the inspection(.
Also, Figure~\ref{fig:waterfall} divides  testing     as follows. Developers perform unit test prior to code complete.  After code complete a standard phase is the integration, which combines program units into workable system ready for system test. Integration,  system test, and acceptance test are often performed by another group.

>>>>

Since 2000, the SEI has been teaching and coaching TSP teams. One of the authors (Nichols) has mentored software development teams and coaches around the world as they deploy TSP within their organizations since 2006.  The  most recent completions were in 2014.
The projects were mostly small to medium, with a median duration of 46 days and a maximum duration of 90 days in major increments. 
Several projects extended for multiple incremental development cycles. 
Median team size was 7 people, with a maximum of 40. See \fig{dist} for the total
effort seen in those projects.

@@

Note that, in that Figure~\ref{fig:waterfall},
several  phases in which the product is created have sub-phases of {\em review} and {\em inspect} to remove defects.
In TSP  reviews,  individuals perform personal reviews of their work products prior to the peer review
(which TSP calls the inspection(.
Also, Figure~\ref{fig:waterfall} divides  testing     as follows. Developers perform unit test prior to code complete.  After code complete a standard phase is the integration, which combines program units into workable system ready for system test. Integration,  system test, and acceptance test are often performed by another group.

### 2015-08-20 phaseDelay.tex
Updates from ShareLaTeX

\conferenceinfo{ICSE}{'16 Austin, Texas}
\title{Live and Let Die? (Delayed Issues   not  Harder to Resolve)}
\numberofauthors{3}
\author{
%\alignauthor 
Tim Menzies \\
       \affaddr{Computer Science}\\
       \affaddr{NC State University, USA}\\
       \email{tim.menzies@gmail.com}
\and%alignauthor 
William Nichols, Forest Shull \\
        \affaddr{Software Engineering Institute}\\
        \affaddr{Carnegie Mellon U, USA}\\
        \email{\{wrn,fjshull\}@sei.cmu.edu} 
\and %\alignauthor 
Lucas Layman \\
       \affaddr{Fraunhofer CESE} \\
       \affaddr{College Park, USA}\\ 
       \email{llayman@fc-md.umd.edu}
} 
\maketitle
\begin{abstract}
Many  practitioners and academics
believe in a delayed issue effect (DIE); i.e.
 as issues linger longer in a system,
 they become exponentially harder to resolve.
This belief
is often  used to justify 
major investments in  new development
processes that promise to retire more issues, sooner.

This paper checks for the delayed issue effect in
171 software projects from around the world (2006 to 2014).
To the best of our knowledge,  this is the largest study
yet published on this effect.
What was observed was the  exact
opposite of the  delayed issue effect. Specifically:
the  time  to resolve 
issues  in one phase  was rarely more than when  issues
were resolved at a later phase.

@@

Also, we note that our development practices have changed in ways that could   mitigate the delayed issued effect.  Previously,
software was large monolithic systems that were write once and maintain
forever. Today, there are more architectures that
support  extensive and faster changes to software. 
For example, 
Microsoft is adjusting its development practices towards a continuous
release paradigm. Theisen et al. reports their experiences where Microsoft architects
continually adjust  their modules in response to security issues~\cite{theisen15}.
Also, upgrades to their Windows operating system is moving from service patches (which occur rarely) to continuous deployment (so there will be no Windows 11- just a stream of continuous updates to what is currently called Windows 10~\cite{bright15}).

@@

The above argument is an anecdotal evidence that  the delayed issue effect might no longer exist. But anecdotes are not really rigorous
evidence. Accordingly,  this article explores the currency of the delayed issue effect.
After some initial definitions, we discuss the
value of checking old ideas. Next, a survey of industrial practitioners and researchers to document a widespread belief that delayed issues have a negative impact on projects.  After that, we  analyze 171 software  projects developed in the period 2006 to 2014 to find  {\em no evidence} of the delayed issue effect:
\bi
\item
To ensure reproducibility,
all the data  used in this study is available in the PROMISE
repository at openscience.us/repo. 
\item
To the best of our knowledge,
this the largest study devoted the delayed issue effect yet conducted.
\ei
Finally, we discuss the validity and implications of our results.

@@

\subsubsection{Comment 2}
Next, we comment on 
requirements errors are the most expensive to fix 
{\em when found during production but the cheapest to fix early in development}. 
Our data from these 171 project does not extend into post-delivery deployment. That said,
there is enough evidence to raise doubts that our projects will exhibit exponentially more
difficult issues after deployment.

@@

We checked for traces of this effect in 171 projects from the period 2006 to 2014.
That data held no trace of the delayed issued effect.
To the best of our knowledge, this paper is the  largest study
of this effect yet performed.

Our explanation for this lack-of-effect is five-fold:
\be
\item The effect might be an historical relic. Evidence:
the effect was first described in the era of punch card computing and non-interactive environments).
\item The effect might have been intermittent (rather than some fundamental law of software). Evidence: we can  found nearly
as many papers reporting the effect~\cite{Boehm76,Boehm81,steck04,Fagan76,Stephenson76} as otherwise~\cite{Royce98,Boehm80,Shull02}.
\item The effect might be confined to very large systems- in which case it would be
acceptable during development to let smaller to medium
sized projects carry some unresolved issues from early phases into later phases.
\item The effect might be mitigated by modern software development approaches that
encourage change and revision of older parts of the system.
\item The effect might be mitigated by modern software development tools
that simplify the process of large-scale reorganization of software systems.
\ee
 %One hypothesized explanation that can be ruled out is that the lack of a phase delay effect is due to using the TSP method. 
Our results beg the question: why did an idea, with so little support, become so widespread in the software engineering literature? No doubt the original evidence was compelling at the time, but much has changed in the realm of software development in the subsequent 40 years. Possibly the concept of the delayed issue effect (or its more specific description: requirements errors are the hardest to fix)
has persisted because, to use Glass's terms on the subject, it seems to be ``just common sense''\cite{glass02}. 
Nevertheless, in a rapidly changing field such as software engineering, even widely-held rules of thumb must be periodically re-verified. 
Progress in the domain of software analytics has made such periodic checks more cost-effective and feasible, and we argue that an examination of local behaviors (rather than simply accepting global heuristics) can be of significant benefit.

>>>>

\conferenceinfo{ICSE}{'16 Austin, Texas}
\title{Live and Let Die? (Delayed Issues   not  Harder to Resolve)}
\numberofauthors{3}
\author{
%\alignauthor 
Tim Menzies \\
       \affaddr{Computer Science}\\
       \affaddr{NC State University, USA}\\
       \email{tim.menzies@gmail.com}
\and%alignauthor 
William Nichols, Forrest Shull \\
        \affaddr{Software Engineering Institute}\\
        \affaddr{Carnegie Mellon U, USA}\\
        \email{\{wrn,fjshull\}@sei.cmu.edu} 
\and %\alignauthor 
Lucas Layman \\
       \affaddr{Fraunhofer CESE} \\
       \affaddr{College Park, USA}\\ 
       \email{llayman@fc-md.umd.edu}
} 
\maketitle
\begin{abstract}
Many  practitioners and academics
believe in a delayed issue effect (DIE); i.e.
 as issues linger longer in a system,
 they become exponentially harder to resolve.
This belief
is often  used to justify 
major investments in  new development
processes that promise to retire more issues, sooner.

This paper tests for the delayed issue effect in
171 software projects conducted in the period from 2006--2014 from around the world.
To the best of our knowledge,  this is the largest study
yet published on this effect.
In that sample, we found the exact
opposite of the  delayed issue effect; i.e.
the  time  to resolve 
issues  in one phase  was rarely more than when  issues
were resolved at a later phase.

@@

Also, we note that our development practices have changed in ways that could   mitigate the delayed issued effect.  Previously,
software was large monolithic systems that were write once and maintain
forever. Today, there are more architectures that
support  extensive and faster changes to software. 
For example, 
Microsoft is adjusting its development practices towards a continuous
release paradigm. Theisen et al. reports their experiences where Microsoft architects
continually adjust  their modules in response to security issues~\cite{Theisen15}.
Also, upgrades to their Windows operating system is moving from service patches (which occur rarely) to continuous deployment (so there will be no Windows 11- just a stream of continuous updates to what is currently called Windows 10~\cite{bright15}).

@@

The above argument is an anecdotal evidence that  the delayed issue effect might no longer exist. But anecdotes are not really rigorous
evidence. Accordingly,  this article explores the currency of the delayed issue effect.
After some initial definitions, we discuss the
value of checking old ideas. Next, a survey of industrial practitioners and researchers to document a widespread belief that delayed issues have a negative impact on projects.  After that, we  analyze 171 software  projects developed in the period 2006--2014 to find  {\em no evidence} of the delayed issue effect:
\bi
\item
To ensure reproducibility,
all the data  used in this study is available in the PROMISE
repository at openscience.us/repo. 
\item
To the best of our knowledge,
this the largest study devoted the delayed issue effect yet conducted.
\ei
Finally, we discuss the validity and implications of our results.

@@

\newpage\subsubsection{Comment 2}
Our data from these 171 project does not extend into post-delivery deployment. That said,
there is enough evidence to raise doubts that our projects will exhibit exponentially more
difficult issues after deployment.

@@

We checked for traces of this effect in 171 projects from the period 2006--2014.
That data held no trace of the delayed issued effect.
To the best of our knowledge, this paper is the  largest study
of this effect yet performed.

Our explanation for this lack-of-effect is five-fold:
\be
\item The effect might be an historical relic. Evidence:
the effect was first described in the era of punch card computing and non-interactive environments.
\item The effect might have been intermittent (rather than some fundamental law of software). Evidence: we can  found nearly
as many papers reporting the effect~\cite{Boehm76,Boehm81,steck04,Fagan76,Stephenson76} as otherwise~\cite{Royce98,Boehm80,Shull02}.
\item The effect might be confined to very large systems- in which case it would be
acceptable during development to let smaller to medium
sized projects carry some unresolved issues from early phases into later phases.
\item The effect might be mitigated by modern software development approaches that
encourage change and revision of older parts of the system.
\item The effect might be mitigated by modern software development tools
that simplify the process of large-scale reorganization of software systems.
\ee
 %One hypothesized explanation that can be ruled out is that the lack of a phase delay effect is due to using the TSP method. 
Our results beg the question: why did an idea, with so little support, become so widespread in the software engineering literature? No doubt the original evidence was compelling at the time, but much has changed in the realm of software development in the subsequent 40 years. Possibly the concept of the delayed issue effect (or its more specific description: requirements errors are the hardest to fix)
has persisted because, to use Glass's terms on the subject, it seems to be ``just common sense''\cite{glass02}. 
Nevertheless, in a rapidly changing field such as software engineering, even widely-held rules of thumb must be periodically re-verified. 
Progress in the domain of software analytics has made such periodic checks more cost-effective and feasible, and we argue that an examination of local behaviors (rather than simply accepting global heuristics) can be of significant benefit.
